---
title: "Deel_Developing Data Products_Week 4 Assignment Pitch"
author: "Jakob Deel"
date: "December 22, 2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  fig.path = "figures/"
  )

setwd("C:/Users/jdeel/Documents/Training/JHU - Coursera/Developing Data Products/Deel Developing Data Products_Week 4 Assignment")
```

## Background

In many places, higher local incomes can be partially offset by higher cost of living, especially housing costs. This can be frustrating for people who move geographies in search of a better financial situation. The dream scenario is to find a location where incomes are higher than usual but higher cost of living does not erase that advantage.

## Study Setup

Locality income and housing costs data is available from the U.S. Census.

My preliminary study uses income and housing cost data for all localities in Virginia, using openly-available  information from the U.S. Census' 2013-2017 American Community Survey 5-year estimates obtained on December 16th, 2018. (https://factfinder.census.gov/faces/tableservices/jsf/pages/productview.xhtml?pid=ACS_17_5YR_S2503&prodType=table).

## Simple Linear Regression

The code below creates a very simple model to predict a locality's median housing cost (in percent relative to statewide mean) using only the locality's median income (again in percent relative to statewide mean). It also calculates a simple example showing that we would expect a locality at the exact statewide income mean to have a median housing cost at around 107% of the statewide mean.

```{r data, include=FALSE}
#reads in data obtained from census.gov
setwd("C:/Users/jdeel/Documents/Training/JHU - Coursera/Developing Data Products/Deel Developing Data Products_Week 4 Assignment")
dat <- read.csv("JD_ACS_17_5YR_S2503_with_ann.csv")

#creates variables for locality median income and housing costs compared to mean
dat$medianincomevsmean <- (dat$medianincome/mean(dat$medianincome))*100
dat$housingmedianvsmean <- (dat$housingmedian/mean(dat$housingmedian))*100
```

```{r model}
#runs simple prediction algorithm on housing vs. mean as function of median income vs. mean
m1 <- lm(dat$housingmedianvsmean~dat$medianincomevsmean)

#calculates example for locality at statewide mean
m1$coefficients[2]*100

```

## Finding advantageous localities

Use the R Shiny app at https://jakobdeel.shinyapps.io/Deel_DDP_Wk4_Project to explore data on the relationship between local income and housing costs and to identify localities below the trendline where extra income will not be absorbed by higher-than-average housing costs
